%
% File created on {2009, 4, 25, 17, 47, 46.104429} by makeDoc.
%
% Section header
\section{The Alpha`Zpol` package } 
\label{label:Alpha`Zpol`}
\alphanote{Zpol}{Documentation revised on August 10, 2004} 
 
\alphausage{Zpol}{The Alpha`Zpol` package contains additional functions for computing
with Z-polyhedra. These functions are expressLatticeWithMat, 
expressZpolWithMat, getMatOfZpol, and getPolOfZpol.}{Alpha/Zpol.m}{Alpha`Zpol`}
\index{Zpol}

\alphanote{Zpol}{The internal representation of a Zpol is:
domain[dim\_Integer, \{\_\_\_String\},
\{\_\_\_zpol[m\_matrix,\{\_\_\_pol\}]\}] (which is m(p)).} 
 
\alphausage{expressLatticeWithMat}{expressLatticeWithMat[sys,m] returns a system where all Z-polyhedra 
using lattices corresponding to the matrix m are rewritten so as 
to appear with matrix m as lattice. Default value of sys is \$result.
expressLatticeWithMat[sys,m,pos] modifies only the 
Z-domain at position pos of the Alpha abstract syntax tree.}{Alpha/Zpol.m}{Alpha`Zpol`}
\index{expressLatticeWithMat}

\alphausage{expressZpolWithMat}{expressZpolWithMat[z,M] attempts to change the representation    
of the Z-polyhedron z. Assuming z = N(P), we expect a result of the 
form w = M(P') where M is the matrix given as second argument, hence 
it returns M(P') where P'=M\^\{-1\} N(P). If M and N do not represent 
the same lattice, it gives a warning and return z unchanged. 
If z is a union of Z-polyhedra, it fails.}{Alpha/Zpol.m}{Alpha`Zpol`}
\index{expressZpolWithMat}

\alphausage{getMatOfZpol}{getMatOfZpol[z] returns the matrix used for identifying the 
lattice of a Z-polyhedron z. If z=M(P), it returns M. If Z is already a 
polyhedron, it returns the identity matrix. If z is a union of 
Z-polyhedra, it fails.}{Alpha/Zpol.m}{Alpha`Zpol`}
\index{getMatOfZpol}

\alphausage{getPolOfZpol}{getPolOfZpol[z] returns the polyhedron used as source for generating 
a Z-polyhedron z. If z=M(P), it returns P. If z is already 
a polyhedron, it returns z. If z is a union of Z-polyhedra, it fails.}{Alpha/Zpol.m}{Alpha`Zpol`}
\index{getPolOfZpol}

